---
title: "03_Shaughnessy_Brianna_2018"
author: "Brianna Shaughnessy"
date: "9/24/2018"
output:
  rmarkdown::html_document:
  theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
library(dplyr)
library(ggplot2)
library(tweenr)
library(gganimate)
library(viridis)
library(transformr)
library(tidyr)
library(readr)
library(wesanderson)
library(devtools)
```

# Homework 3: Data Visualization

## 1: Problems from Whitlock and Schluter

####10a. 95% Confidence Interval for the population mean
6.86 < true mean < 9.76

####10b. Provide an interpretation of the 95% Confidence Interval
This 95% Confidence Interval indicates that we are 95% certain that the true value of the mean population falls within the calculated range between 6.9 and 7.8.

```{r}
Genes <- read.csv("GeneRegulation.csv")
GenesCI <- Genes%>%
  #find the average/mean of the population
  summarize( avg_gene = mean(ngenes),
   #find the standard deviation of the mean         
             sd_gene = sd(ngenes),
   #calculate the standard error
            std_err = sd_gene/(sqrt(109)),
  #find the 95% Confidence Interval
            CI_lower = avg_gene - (2*std_err),
            CI_upper = avg_gene + (2*std_err))
GenesCI
```

####17. Is the Confidence Interval interpreted correctly? Explain.
This is not the correct interpretation of the confidence interval because it is not a measure of probability. 

####18. Corpseflowers
**Mean:** 70.1  
**Standard Deviation:** 48.5  
**Standard Error:** 15.3  
**95% Confidence Interval:** Lower: 39.4 < true mean < 100.8  
**d.** With a larger sample size the mean would remain within the same range as this smaller sample.
**e.** Assuming a normal districution: increasing sample size decreases standard deviation - within the equation for standard deviation we are dividing a calculated number by the sample size. Thus: dividing by a larger number will produce a smaller product but under certain circumstances there is a threshold where the sample mean (x) will approach the population mean and the two will even eachother out. 
**f.** Increasing sample size increases precision in our sampling distribution. I would expect standard error to decrease with increasing sample size.

```{r}
getwd()
Flowers <- read.csv("Corpseflowers.csv")
Flower_Stats <- Flowers %>%
  summarize(avg_flowers = mean(numberOfBeetles),
            sd_flowers = sd(numberOfBeetles),
            SE_flowers = sd_flowers/(sqrt(10)),
            Upper_Quartile = avg_flowers + (2*SE_flowers),
            Lower_Quartile = avg_flowers - (2*SE_flowers))
Flower_Stats
```

## 2: Arctic Sea Ice 1978 - 2016
####2.1 Load the data and make month names into factors. Are they in the right order?
Not all months are in the correct order. This may be because it looks like the data was collected during different years.

```{r}
Sea_Ice <- read_csv("NH_Sea_Ice.csv")
Sea_Ice_Data <-Sea_Ice %>%
  mutate(Month_Name = factor(Month_Name)) %>%
  mutate(Month_Name = forcats::fct_inorder(Month_Name))
  
levels(Sea_Ice_Data$Month_Name)
?levels
```

####2.2 What is the order of factor levels that result?

**fct_rev()**reverses the order of factor levels, so on this dataset this function outputs the factor levels from the above question but in reverse order.  
**fct_relevel()**allows us to move levels around. With my code I moved August, February, and January to see that whatever levels are listed will appear in the order that is represented in my code.  
**fct_recode()**allows the user to change levels by hand. For my code I replaced "Feb" with "Month"  


```{r}
levels(fct_rev(Sea_Ice_Data$Month_Name))
?fct_rev
levels(fct_relevel(Sea_Ice_Data$Month_Name, "Aug", "Feb", "Jan"))
?fct_relevel
levels(fct_recode(Sea_Ice_Data$Month_Name, Month = "Feb"))
?fct_recode
```

Mutate month name to get months in the right order, from January to December. Show that it worked with levels()

```{r}
Sea_Ice_Data <- Sea_Ice_Data %>%
  mutate(Month_Name = fct_relevel(Sea_Ice_Data$Month_Name,
                                  "Jan", "Feb", "Mar", "Apr", "May",
                                  "Jun", "Jul", "Aug", "Sep", "Oct",
                                  "Nov", "Dec"))
levels(Sea_Ice_Data$Month_Name)
```


####2.3 Make a column called Season that is a copy of Month Name and turn it into a factor vector with the levels Winter, Spring, Summer, Fall in that order. Use levels() on Sea_Ice_Data$Season to show that it worked.

```{r}
#Make a column called Season that is a copy of Month_Name.
Sea_Ice_Data <- Sea_Ice_Data %>%
  mutate(Season = fct_recode(Month_Name, Winter = "Jan", Winter =
                               "Feb", Spring = "Mar", Spring =
                               "Apr", Spring = "May", Summer = 
                               "Jun", Summer = "Jul", Summer = 
                               "Aug", Fall = "Sep", Fall =
                               "Oct", Fall = "Nov", Winter =
                               "Dec")) 
                             #Give the factors levels for Season.
#Show the levels
levels(Sea_Ice_Data$Season)
```

####2.4 Make a boxplot showing the variability in sea ice extent every month.

```{r}
Box_Plot <- ggplot(data = Sea_Ice_Data,
                   mapping = aes(x = Month_Name, y = Extent))+
  geom_boxplot()
Box_Plot
```


####2.4(cont) Use dplyr to get the annual minimum sea ice extent. Plot minimum ice by year, and add a trendline (either smooth spline or a straight line)

```{r}

#Find the annual minimum by targetting the Year column and summarizing the minimum of Extent grouped by year
Ice_Stats <- Sea_Ice_Data %>%
  group_by(Year) %>%
  summarize(minimum = min(Extent))

#Make the plot

ggplot(Ice_Stats,
       aes(x = Year, y = minimum)) +
      geom_point()+
      stat_smooth(method = lm)
```

####2.5 With the original data, plot sea ice by year, with different lines for different months. Then, use facet_wrap and cut_interval(Month, n=4) to split the plot into seasons.
  Using n = 4 just split up months by 4 but not in order of Season - so I facet-wrapped the Season column instead.

```{r}

graph_by_year <- ggplot(data = Sea_Ice_Data,
                mapping = aes(x=Year, y=Extent, 
                              color = Month, group = Month)) +
#add the points
              geom_line()
#Now facet wrap to add Seasons
graph_by_year + facet_wrap(~Season)
```


####2.6) Last, make a line plot of sea ice by month with different lines as different years. Color by year, a different theme, and whatever other annotations, changes to axes, etc. 

```{r}
graph_by_month <- ggplot(data = Sea_Ice_Data,
                         mapping = aes(x = Month_Name, y = Extent, 
                                       group = Year, color = Year)) +
  geom_line()+
  scale_color_gradientn(colors = wes_palette("Zissou1"))+
  guides(colors = "none") +
  theme_dark()+
  labs(x = "Month",
       y = "Sea Ice Extent",
       title = "Arctic Sea Ice 1978-2016")
  
Arctic_Ice <- graph_by_month + theme(panel.grid.major = element_blank())
Arctic_Ice
```

```{r}
Arctic_Ice +
  transition_time(Year)
Arctic_Ice +
  transition_reveal(Year, along = Year)
```